---
title: "General Social Survey"
author: "Dana Hoffman"
date: "10/24/19"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


\(H_0\):There is no difference in the people who agree on both the death penalty and the right to own guns.

\(H_A\):There is a statistical difference between the people who agree on the death penalty and the right to own guns.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We will be using both graphical descriptive and numerical descriptive results.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


### Descriptive Results

The result concluded that there is a statistical difference between the people who are in favor of the death penalty and the people in favor of gun rights.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~CAPPUN + OWNGUN,data=mydata4)
barchartGC(~CAPPUN + OWNGUN,data=mydata4,type="percent")
```

 In these charts you can see that the amount of people in favor of the death penalty is almost equivalent to the amount of people against it. While the amount of people who oppose the right to own guns is significantly higher than those in favor.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=mydata4)
rowPerc(table1)
colPerc(table1)
```

This table shows that 48% of the people were in favor of the death penalty while 52% was against it. It also shows that 29% of the people are against the right to own a gun while 71% percent are in favor.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```
Here we get a P-value of 0.0202, we reject the null hypotheses.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

\(H_0\):There is no difference in the people who agree on both the bigbang theory and the belief in the afterlife

\(H_A\):There is a statistical difference between the people who agree on the bigbang theory and the belief in the afterlife.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We will be using both graphical descriptive and numerical descriptive results.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

The result concluded that there is a statistical difference between the people who are in favor of the big bang theory and the independent belief of the afterlife.


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~BIGBANG + POSTLIFE,data=mydata4)
barchartGC(~BIGBANG + POSTLIFE,data=mydata4,type="percent")
```

 In these charts you can see that the amount of people in favor of the big bang theory is significantly higher to the amount of people against it. While the amount of people who oppose the right to is significantly lower than those in favor of the big bang theory and independent of belief in the afterlife. Basically it states that the big bang theory is true if they believe the afterlife is false.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~BIGBANG + POSTLIFE, data=mydata4)
rowPerc(table1)
colPerc(table1)
```
This table shows that 82% of the people believe in bigbang theory and 19% people who oppose it.  It also shows that 92% of the people are against the bigbang theory while 8% percent are in favor.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```
Here we get a P-value of 0.0801, we fail to reject the null hypotheses.
# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a person’s opinion on death penalty (CAPPUN) is independent of their race (RACE).

\(H_0\):There is no difference in the people who agree on both the death penalty and is independent of their race.

\(H_A\):There is a statistical difference between the people who agree on the death penalty and is independent of their race.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

We will be using both graphical descriptive and numerical descriptive results.

### Descriptive Results

The result concluded that there is a statistical difference between the people who are in favor of the death penalty and is independent of their race.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

We use bar charts to compare Numerical vs Numerical data.

```{r}
barchartGC(~CAPPUN + RACE,data=mydata4)
barchartGC(~CAPPUN + RACE,data=mydata4,type="percent")
```
 It states that the majority of white people believe in the death penalty. While the amount of other races who oppose and favor the death penalty are significantly lower the white individuals who voted.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~CAPPUN + RACE, data=mydata4)
rowPerc(table1)
colPerc(table1)
```
According to these stats most white gun owners are also in favor of the death penalty. Statistically the data states if you own a gun you’re in favor of the death penalty

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
```
Here we get a P-value of 0.0009, we reject the null hypotheses.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion to all three of the questions the first and last question due to the P-value having a value of less than 5% we reject the null hypothesis. However, question two we reject to fail the null hypothesis because there value is more than 5%.
In question one The result concluded that there is a statistical difference between the people who are in favor of the death penalty and the people in favor of gun rights.
In question two it concluded that Basically it states that the big bang theory is true if they believe the afterlife is false.
Lastly, question three According to these stats most white gun owners are also in favor of the death penalty. Statistically the data states if you own a gun you’re in favor of the death penalty


